---
title: "Timeseries HW1"
author: "Ze YANG"
date: "September 10, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\section{Problem 1}

## 1.a

\begin{equation}
\begin{split}
  \mathbb{E}\left[X_t\right] &= \mathbb{E}\left[a+bZ_t+cZ_{t-1}\right] = a \\
  \mathrm{\mathbb{C}ov}\left[X_t, X_{t+h}\right] &= \mathrm{\mathbb{C}ov}\left[a+bZ_t+cZ_{t-1}, a+bZ_{t+h}+cZ_{t+h-1}\right] \\
  &= b^2 \mathrm{\mathbb{C}ov}\left[Z_t, Z_{t+h}\right] + c^2 \mathrm{\mathbb{C}ov}\left[Z_{t-1}, Z_{t+h-1}\right]\\
  &\quad + bc \mathrm{\mathbb{C}ov}\left[Z_{t+h}, Z_{t-1}\right] + bc \mathrm{\mathbb{C}ov}\left[Z_t, Z_{t+h-1}\right] \\
  &= \begin{cases}
  (b^2+c^2) \sigma^2 & h=0\\
  bc\sigma^2 & h=\pm 1\\
  0 & \text{otherwise}
  \end{cases}\\
  \mathbb{E}\left[X_t^2\right] &= \mathrm{\mathbb{V}ar}\left[X_t\right] + \mathbb{E}\left[X_t\right]^2 = a^2 + (b^2+c^2)\sigma^2 < \infty
\end{split}
 \end{equation} 
 Therefore, $X_t$ is stationary.

## 1.b

 \begin{equation}
   \begin{split}
     \mathbb{E}\left[X_t\right] &= a + \sigma^2\\
     \mathrm{\mathbb{C}ov}\left[X_t, X_{t+h}\right] &= \mathrm{\mathbb{C}ov}\left[a+X_t^2, a+X_{t+h}^2\right] \\
     &= \mathrm{\mathbb{C}ov}\left[X_t^2, X_{t+h}^2\right]\\
     &=\begin{cases}
     2\sigma^4 & h=0\\
     0 & \text{otherwise}
     \end{cases} \\
     \mathbb{E}\left[X_t^2\right] < \infty
   \end{split}
 \end{equation}
  Therefore, $X_t$ is stationary.

## 1.c

 \begin{equation}
   \begin{split}
     \mathbb{E}\left[X_t\right] &= \mathbb{E}\left[Z_tZ_{t-1}\right] = 0\\
     \mathrm{\mathbb{C}ov}\left[X_t, X_{t+h}\right] &= \mathrm{\mathbb{C}ov}\left[Z_tZ_{t-1}, Z_{t+h}Z_{t+h-1}\right] \\
     &= \mathbb{E}\left[Z_tZ_{t-1}Z_{t+h}Z_{t+h-1}\right] - \mathbb{E}\left[Z_tZ_{t-1}\right] \mathbb{E}\left[Z_{t+h}Z_{t+h-1}\right]\\
     &=\mathbb{E}\left[Z_tZ_{t-1}Z_{t+h}Z_{t+h-1}\right]\\
     &=\begin{cases}
     \mathbb{E}\left[Z_t^2\right]\mathbb{E}\left[Z_{t-1}^2\right] & h=0\\
     0 & \text{otherwise}
     \end{cases} \\
          &=\begin{cases}
     \sigma^4 & h=0\\
     0 & \text{otherwise}
     \end{cases} \\
     \mathbb{E}\left[X_t^2\right] < \infty
   \end{split}
 \end{equation}
 Therefore, $X_t$ is stationary.

\section{Problem 2}
Since $\{\epsilon_t\}$ is i.i.d. white noise, let $\mathrm{\mathbb{V}ar}\left[\epsilon_t\right] = \sigma^2$.
\begin{equation}
  \mathbb{E}\left[X_t\right] = \mathbb{E}\left[\epsilon_t + \epsilon_{t-1} \epsilon_{t-2} \right] = 0
\end{equation}
\begin{equation}
  \begin{split}
    \mathrm{\mathbb{C}ov}\left[X_t, X_{t+h}\right] &= \mathrm{\mathbb{C}ov}\left[\epsilon_t + \epsilon_{t-1} \epsilon_{t-2}, \epsilon_{t+h}+ \epsilon_{t+h-1} \epsilon_{t+h-2}\right] \\
  &=\mathbb{E}\left[(\epsilon_t + \epsilon_{t-1} \epsilon_{t-2})(\epsilon_{t+h}+ \epsilon_{t+h-1} \epsilon_{t+h-2})\right] \\
  &= \mathbb{E}\left[\epsilon_t \epsilon_{t+h}+ \epsilon_{t-1} \epsilon_{t-2} \epsilon_{t+h} + \epsilon_{t+h-1} \epsilon_{t+h-2} \epsilon_t + \epsilon_{t-1} \epsilon_{t-2}\epsilon_{t+h-1} \epsilon_{t+h-2}\right]\\
  &= \begin{cases}
  \mathbb{E}\left[\epsilon_t^2 + \epsilon_{t-1}^2 \epsilon_{t-2}^2\right] & h=0\\
  0 & |h|>0
  \end{cases}\\
  &= \begin{cases}\sigma^2 + \sigma^4 & h=0\\
  0 & |h|>0
  \end{cases}
  \end{split}
\end{equation}
Therefore by definition, $\{X_t\}$ is weak white noise. But
\begin{equation}
  \mathbb{E}\left[X_t | \mathcal{F}_{t-1}\right] = \mathbb{E}\left[\epsilon_t | \mathcal{F}_{t-1}\right] + \epsilon_{t-1} \epsilon_{t-2} = \epsilon_{t-1} \epsilon_{t-2} \ne 0
\end{equation}
Therefore, $\{X_t\}$ is not MDS.

\section{Problem 3}
\begin{equation}
  \begin{split}
    \mathrm{\mathbb{C}ov}\left[X_t, X_{t+h}\right] &= \mathrm{\mathbb{C}ov}\left[Y_t - Y_{t-1}, Y_{t+h}-Y_{t+h-1}\right] \\
    &= \mathrm{\mathbb{C}ov}\left[Y_t, Y_{t+h}\right] - \mathrm{\mathbb{C}ov}\left[Y_{t-1}, Y_{t+h}\right] - \mathrm{\mathbb{C}ov}\left[Y_t, Y_{t+h-1}\right] + \mathrm{\mathbb{C}ov}\left[Y_{t-1}, Y_{t+h-1}\right] \\
    &= \gamma_Y(h) - \gamma_Y(h+1) - \gamma_Y(h-1) + \gamma_Y(h) \\
    &= 2\gamma_Y(h)- \gamma_Y(h+1) - \gamma_Y(h-1)\\
    &=: \gamma_X(h)
  \end{split}
\end{equation}
which is a function of $h$, independent of $t$. Moreover, $\mathbb{E}\left[X_t\right]=0$. Therefore, $X_t$ is also stationary.


# Problem 4
## 4.1
```{r}
library(quantmod)
goog = getSymbols('GOOG',from='2017-1-1',to='2017-12-31',auto.assign=F)
logret = dailyReturn(Ad(goog), type='log')
plot(logret, main='Log return of GOOG, Year 2017')
```
The log returns exhibit centain level of volatility clustering. However, there is no clear evidence of non-stationarity by just looking at the plot.

## 4.2
```{r}
acf(logret, main='ACF')
```

## 4.3
```{r}
Box.test(logret, lag=20, type="Ljung-Box")
```
The p-value is large. We fail to reject $H_0: \{\rho_X(h) = 0$ for some $h\}$. That is, there is no evidence of serial correlation up to lag 20.
