---
title: "Time Series HW3"
author: "Ze YANG"
date: "9/25/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Ruppert & Matteson
## Problem 1


```{r fig.width=8, fig.asp=1}
# Load data
library(tseries)
library(forecast)
library(Ecdat)
data(IncomeUK)
consumption = IncomeUK[,2]
par(mfrow=c(3,2))
plot(consumption)
acf(diff(log(consumption), 1), 30)
plot(diff(log(consumption), 4, 1))
acf(diff(log(consumption), 4, 1), 30)
plot(diff(diff(log(consumption), 4, 1), 1))
acf(diff(diff(log(consumption), 4, 1), 1), 30)
```

**The behavior of consumption**

1. The size of seasonal oscillations seem to *increase* with time.
2. There is an upward trend (hence nonstationarity).
3. There is an evident seasonal component. The ACF of the first-difference suggests the seasonality has a period = 4.
4. After applying lag-4 difference, the series still has a trend.
5. After applying the first difference of the lag-4 difference, the series appears to be nice.

Behavior 1 suggests we apply a Box-Cox (log) transformation to stablize the size of seasonal oscillations. Behavior 2-5 implies we apply both seasonal and non-seasonal differencing to remove two types of serial dependencies.

## Problem 2
```{r}
fit.res = arima(log(consumption), 
                order=c(1,1,1), 
                seasonal=list(order=c(1,1,0), period=4))
summary(fit.res)
plot(log(consumption))
lines(fitted(fit.res), col='red')
legend(1971.5, 10.8, 
       legend=c("True", "Fitted"), 
       col=c("black", "red"),
       lty=c(1,1), cex=0.8)
```

An $ARIMA((1,1,1) \times (1,1,0)_4)$ model gives a good fit to the data.

## Problem 3
```{r}
acf(fit.res$residuals, 20)
Box.test(fit.res$residuals, fitdf=3, lag=20, type="Ljung-Box")
```

- The residual ACF displays no evidence of correlation till lag 20. There is no residual autocorrelation detected.

- The p-value of Ljung-Box tests for the residuals is very big. Fail to reject null: no residual autocorrelation.

# Three-Month Treasury Bills
## (a)
```{r}
tbill.full = read.csv('TB3MS.csv')
tbill = tbill.full[(as.Date(tbill.full$DATE) <= '2012-12-31') & (
  as.Date(tbill.full$DATE) >= '1990-1-1'),]
tbill = ts(tbill$TB3MS)
plot(tbill)
```

- A clear non-mean-reverting trend can be observed. The orginal series is not stationary.

## (b)
```{r fig.width=8, fig.asp=0.5}
par(mfrow=c(1,2))
plot(diff(tbill, differences=1))
adf.test(diff(tbill, differences=1))
```

We choose to use the First difference of the log-transformed series. The ADF test yield 0.03763 p-value, indicates a rejection of the null hypothesis ($H_0$: the series is integrated) at 95% confidence level. And the plot of the series seems to be approximately stationary indeed.

## (c)
```{r}
fit.tbill = auto.arima(
  tbill, d=1, D=1, max.p=10, max.q=10, 
  ic='aicc', approx=F, step=F)
summary(fit.tbill)
```

The selected model is $ARIMA(3,1,0)$.

## (d)
```{r fig.width=8, fig.asp=1}
par(mfrow=c(2,2))
plot(fit.tbill$residuals)
acf(fit.tbill$residuals)
plot(tbill)
points(fitted(fit.tbill), col='red')
legend(0, 15.0, 
       legend=c("True", "Fitted"), 
       col=c("black", "red"),
       lty=c(1, NA), pch=c(NA, 1), cex=0.8)
qqnorm(fit.tbill$residuals)
qqline(fit.tbill$residuals)
```

The model seems to be a reasonable fit. From upper left to bottom right are: plot of the residuals, ACF of the residuals, true V.S. fitted in-sample, Q-Q plot of the residuals. There is no significant autocorrelation in residuals, and the in-sample fit seems to be good.


## (e)
```{r}
pred.res = predict(fit.tbill, n.ahead = 3)
plot(tbill)
lines(pred.res$pred, col='red')
lines(pred.res$pred+1.96*pred.res$se, col='red', lty=2)
lines(pred.res$pred-1.96*pred.res$se, col='red', lty=2)
```

## (f)
```{r}
upper = pred.res$pred+1.96*pred.res$se
lower = pred.res$pred-1.96*pred.res$se
for(i in 1:3) {
  print(paste('Center:', pred.res$pred[i], 
              '95% CI: [', upper[i], lower[i], ']'))
}
```
The point forecast and 95% CIs of the subsequent 3 months are printed as above.

```{r}
tbill.test = tbill.full[(as.Date(tbill.full$DATE) > '2012-12-31') & (
  as.Date(tbill.full$DATE) <= '2013-3-30'),]
tbill.test
```

We conclude that all the three out-of-sample true values are contained in the CIs.