---
title: "Timeseries HW5"
author: "Ze YANG"
date: "10/7/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Macro Vars
## Question (a)

```{r fig.width=8, fig.asp=1}
library(vars)
source("https://people.orie.cornell.edu/davidr/SDAFE2/Rscripts/SDAFE2.R")
macro.data = read.csv('MacroVars.csv')
acf(macro.data)
```

```{r fig.width=8, fig.asp=1}
par(mfrow = c(3,3))
ccf(macro.data$GDP251, macro.data$GDP251)
ccf(macro.data$GDP251, macro.data$FYFF)
ccf(macro.data$GDP251, macro.data$CPIAUCSL)
ccf(macro.data$FYFF, macro.data$GDP251)
ccf(macro.data$FYFF, macro.data$FYFF)
ccf(macro.data$FYFF, macro.data$CPIAUCSL)
ccf(macro.data$CPIAUCSL, macro.data$GDP251)
ccf(macro.data$CPIAUCSL, macro.data$FYFF)
ccf(macro.data$CPIAUCSL, macro.data$CPIAUCSL)
```
**Interesting Observations**

- All the three series have autocorrelation on their own.
- Some pairs of series also have cross serial correlation. The most significant one is the cross-serial correlation between `GDP251` and `FYFF`. The second ccf plot suggests there is positive correlations on negative lags, and negative correlations on positive lags for this pair. 
- That is: An above-average change in GDP predicts an above-average change in federal funds rate. But an above-average change in Federal funds rate redicts a below-average change in GDP. This is kinda consistant to the debt-GDP cycle in the real life.

## Question (b)
```{r fig.width=8, fig.asp=1}
var.fit = VAR(macro.data, lag.max=8)
var.fit$p
summary(var.fit)
```

A model of $p=7$ is chosen in this cased based on AIC criterion.

## Question (c)
```{r fig.width=8, fig.asp=1}
acf(residuals(var.fit))
mLjungBox(residuals(var.fit), lag=20)
```

**Comments**

- The multivariate Ljung-Box test has a p-value = 0.971. Meaning that we fail to reject the null hypothesis: the residual series has no serial correlation within or between all variables.
- The ACF plots of residuals also display no significant serial correlation for the residuals. Thus we conclude that the VAR fit is likely to be adequate.

# Ruppert & Matteson R Lab
## Problem 7
```{r}
TbGdpPi = read.csv("TbGdpPi.csv", header=TRUE)
TbPi.diff = ts(apply(TbGdpPi[,-2], 2, diff), start=c(1955, 2), freq=4)
plot(TbPi.diff)
```

```{r fig.width=8, fig.asp=0.8}
acf(TbPi.diff^2)
mLjungBox(TbPi.diff^2, lag=8)
```

**Answer**:
Yes, the joint series exhibit conditional heteroskedasticity. Because both series has clear volatility clustering pattern: $r_t$ has high variance in early 1980s, while $\pi_t$ has high variance during 2000s.

## Problem 8
```{r fig.width=8, fig.asp=0.8}
EWMA.param = est.ewma(lambda.0 = 0.95, innov=TbPi.diff)
EWMA.param$lambda.hat
EWMA.Sigma = sigma.ewma(lambda = EWMA.param$lambda.hat, innov=TbPi.diff)

par(mfrow = c(2,2))
plot(ts(EWMA.Sigma[1,1,]^.5, start = c(1955, 2), frequency = 4),
     type = 'l', xlab = "year", ylab = NULL,
     main = expression(paste("(a) ", {hat(sigma)[{"1,t"}]})))
plot(ts(EWMA.Sigma[1,2,], start = c(1955, 2), frequency = 4),
     type = 'l', xlab = "year", ylab = NULL,
     main = expression(paste("(b) ", {hat(sigma)[{"12,t"}]})))
plot(ts(EWMA.Sigma[1,2,]/(sqrt(EWMA.Sigma[1,1,]* EWMA.Sigma[2,2,])),
        start = c(1955, 2), frequency = 4),type = 'l', xlab = "year", 
     ylab = NULL,main = expression(paste("(c) ", {hat(rho)[{"12,t"}]})))
plot(ts(EWMA.Sigma[2,2,]^.5, start = c(1955, 2), frequency = 4),
     type = 'l', xlab = "year", ylab = NULL,
     main = expression(paste("(d) ", {hat(sigma)[{"2,t"}]})))
```

**Answer**: The estimated persisitence parameter is $\hat{\lambda} = 0.8665812$.

## Problem 9
```{r fig.width=8, fig.asp=0.8}
n = dim(TbPi.diff)[1]
d = dim(TbPi.diff)[2]
stdResid.EWMA = matrix(0,n,d)
for(t in 1:n){
  stdResid.EWMA[t,] = TbPi.diff[t,] %*% matrix.sqrt.inv(EWMA.Sigma[,,t])}
mLjungBox(stdResid.EWMA^2, lag=8)
acf(stdResid.EWMA^2)
```

**Answer**: It is likely that the EWMA model is not very adequate based on the results of multivariate Ljung-Box test. Because:

- The p-value of the test is $0.002$, small. Therefore we can reject the null hypothesis $H_0: \{\text{The standardized residual series have neither serial correlation within series nor between series up to lag 8}\}$. 
- The ACF plots of the standardized residual series also suggests that there are significant amount of remaining serial correlations (the parts beyond the dashed blue line thresholds).