---
title: "Timeseries HW6"
author: "Ze YANG"
date: "10/15/2018"
output: pdf_document
---

# Ruppert Page 397 Problem 9-11

## Problem 9

```{r}
library(car)
library(AER)
library(forecast)
library(arfima)
data("USMacroG")
MacroDiff = as.data.frame(apply(USMacroG, 2, diff))
attach(MacroDiff)
fit1 = arima(unemp, order=c(1,0,0), xreg=cbind(invest, government))
```
```{r}
fit.lm = lm(unemp ~ invest + government)
c(AIC(fit1), AIC(fit.lm))
par(mfrow=c(1,2))
acf(residuals(fit.lm))
acf(residuals(fit1))
durbinWatsonTest(fit.lm, 5)
```

**Comments:** 

- The AR(1) model has smaller AIC (86.85233 V.S. 138.88245).
- The ACF plot shows that the residuals of AR(1) fit has no significant serial correlation. Hence AR(1) model is a better fit.
- The Durbin-Watson test on the simple linear regression model suggests we reject the null hypothesis for lag = 1 and 2, that is, the residuals exhibits serial correlation for lag = 1,2. 


## Problem 11
```{r}
fit2 = arima(unemp, order=c(2,0,0), xreg=cbind(invest, government))
fit11 = arima(unemp, order=c(1,0,1), xreg=cbind(invest, government))
c(AIC(fit2), AIC(fit11), AIC(fit1), AIC(fit.lm))
```
The AR(2) model has a smaller AIC than AR(1) and ARMA(1,1).


# Ford Data
```{r}
library(quantmod)
FORD = getSymbols('F', from='1995-1-1', to='2017-12-31', auto.assign = F)
volume = FORD$F.Volume
logvolume = log(volume)
par(mfrow=c(1,2))
plot(logvolume)
acf(logvolume, 100)
```

- Yes, it is probably a good idea to fit ARFIMA model on the log volume data, because it as an ACF plot that decays very slowly, which matches the long-memory characteristic that ARFIMA models.



```{r, eval=FALSE}
# Takes very long time to run
maxp = 5
maxq = 5
holdaic = matrix(0, nrow=(maxp+1), ncol=(maxq+1))
for (p in 0:maxp) {
  for (q in 0:maxq) {
    holdaic[(p+1), (q+1)] = summary(
      arfima(ts(logvolume), order=c(p,0,q), quiet=TRUE))$aic
    print(paste('Fitted model', p, q))
  }
}
```

```{r}
# I saved the results in a file
holdaic = as.data.frame(read.csv('holdaic.csv'))
holdaic
```

```{r}
fit53 = arfima(ts(logvolume), order=c(5,0,3), quiet=TRUE)
```

```{r}
coef(fit53)
```

```{r}
AIC(fit53)
```

The best model based on AIC is ARFIMA(5, 0.4966, 3), which has an AIC of -12020.64.