---
title: "Homework 6"
author: "Ze Yang (zey@andrew.cmu.edu)"
date: "Due Thursday, December 14 at 3:00 PM"
output: pdf_document
---

\large

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod)
library(tseries)
```

You should complete this homework solely within R Markdown.
You do not need to show any derivations that were required
to complete Question 1, just implement the analysis in R.

\vspace{.2in}

__Question 1__ 

Assume that $X_1, X_2, \ldots, X_n$ are iid from the Poisson$(\lambda)$
distribution.

Assume you want to test $H_0\!: \lambda = 4.3$ versus $H_1\!: \lambda < 4.3$
using the Wald test.
Calculate the p-value when the data are those found via 

```{r}
x = read.table("http://www.stat.cmu.edu/~cschafer/MSCF/PoisHypTest.txt")
x = x$V1
# use the poisson mle 
lambda_hat = mean(x)
lambda_0 = 4.3
se = sqrt(lambda_0 / length(x))
# test statistic
wald_t = (lambda_hat - lambda_0)/se

# reject H0 if T < -z_alpha
alpha = c(0.001, 0.05, 0.1)
z = qnorm(1-alpha)
wald_t < -z
```
Therefore, we fail to reject $H_0$ for $\alpha \leq 0.1$. So the p-value is greater than 0.1. The actual p-value is calculated as

```{r}
p_val = pnorm(wald_t)
# sanity check: reject H0 if alpha > p, FTR if a < p
test_alpha = c(0, p_val - 10e-6, p_val + 10e-6, 1)
test_z = qnorm(1-test_alpha)
wald_t < -test_z
print(p_val)
```
Hence the p-value is $0.3515102$.

\vspace{.2in}

__Question 2__

The __Augmented Dickey-Fuller (ADF) Test__ is a test utilized
in time series analysis in order to assess stationarity.
There is a function `adf.test()` in the package `tseries`
which implements this test.
Take a look at `help(adf.test)` for some details.

a. If our objective is to see if there is strong evidence
that a time series is stationary, how should the
argument `alternative` be set when using `adf.test()`?

b. Consider the following R commands. This will read in the
stock data discussed in lecture, and then run the ADF test
on the first stock in the sample. Use this as a starting point
to write code to loop over all 1000 stocks and get the p-values
for each test. How many of the tests have p-values less than 0.05?

c. Run the p-values found in part (b) through the Benjamini-Hochberg
procedure described in lecture. Are any of the series found to be
stationary using this approach (again using $\alpha=0.05$)? Comment on the reason(s) for any differences found.


- __Answer (a)__: Since the hypothesis tests are formulated to test __against__ the null hypothesis, for our purpose, we want to find stationary stocks. So we should set $H_0:$ Not stationary, i.e. $H_1:$ stationary. We will use `alternative='s'`.


```{r results='hide', message=FALSE, warning=FALSE}
stocksample = read.table("stocksample.txt", header=T,
                          sep="\t", comment.char="")
p_values = data.frame(id=rep(0,1000), pval=rep(0,1000))
for (i in 1:1000) {
  adf = adf.test(as.numeric(log(stocksample[i,5:34])), alternative="s")
  p_values$id[i] = i
  p_values$pval[i] = adf$p.value
}

p_values = p_values[order(p_values$pval),]
```
```{r}
str(p_values)
nrow(p_values[p_values$pval < 0.05,])
```
So there are 61 tests whose p-value is smaller than 0.05, 6.1% of the total.

```{r}
sum(p.adjust(p_values$pval, method='BH') < 0.05)
```

After the adjustment with the Benjamini-Hochberg procedure, there is no p-value that is smaller than 0.05.

__Comments__

- The different is due to the multiple testing problem. Assume the null hypothesis is true (stock prices are **NOT** stationary), then the p-value follows a uniform distribution. If we sequentially do many tests, we will ultimately see some rejections, i.e. small p-values. 

- If we naively apply some constant confidence level, say $\alpha = 0.05$, and regard the tests with $p < 0.05$ as rejects, we are likely to make false positive rejections.

- The Benjamini-Hochberg procedure adjusts the p-values by controling the false positive rates. So it rules out our false discoveries, and we end up seeing zero test having small p-values after this adjustment.

- It's likely that no stock price series is truely stationary.
