---
title: "Homework 2"
author: "Ze Yang (zey@andrew.cmu.edu)"
date: "Due Thursday, September 14 at 3:00 PM"
output: pdf_document
---

\large

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You should submit the Rmd file with your answers in the
appropriate spaces. Rename the file as `YOURANDREWID_HW2.Rmd`
and submit it via Canvas. Also submit the `.pdf` file that
is produced.

Of course, any code that is written should be tested. In each
case you are requested to include some __simple__ examples showing that
your code works. Think about how your example(s) can illustrate 
the range of possibilities that the code could face.

\vspace{.2in}

1. Suppose that `charvec` is a vector of type `character`.
Write a __single__ line of R code that returns a new, single character
string that consists of all of the elements of `charvec`
concatenated end to end. For example,
`charvec = c("a","b","c")`
should become `"abc"`.

**Solution**:
```{r}
concat.str = function(str.vec) {
  #' Concatenate vector of strings into a new string.
  #' @param str.vec: a vector of strings.
  #' @return the concatenated string.
  return(paste(str.vec[!is.na(str.vec)], collapse = ''))
}
```
**Testcases**:
```{r}
# vanilla case, "Hello world!" expected
stopifnot(concat.str(c("Hello"," ","world!")) == 
            "Hello world!")

# we should not include NA to the concatenated string,
# if there is any. "abc ef g" expected
stopifnot(concat.str(
  c("a",NA, "b","c", NA, " ", "e", "f", " ", "g")) == 
    "abc ef g")

# when is.na() is applied to empty vector, it will give a warning
# but we can live with that, it's natural expected behavior.
# Note that concat.str returns an empty string.
stopifnot(concat.str(c()) == "")

# when trying to concatenate an NA-vector, also return empty string
stopifnot(concat.str(c(NA, NA, NA)) == "")

# Work on integer vectors for free, "20171211" expected.
stopifnot(concat.str(c(2017, 12, 11)) == "20171211")
```


2. Write an R function that takes a numeric vector as input
and returns the ``five-number summary`` as a __list__.
The five-number summary consists of the minimum, the
25th, 50th, and 75th percentile, along with the mapximum.
Name the components of the list appropriately.

**Solution**:
```{r}
percentile.list = function(vec) {
  #' Calculate the five-number summary of a numeric vector.
  #' @param vec: a numeric vector.
  #' @return list of 5-number summary.
  l.percentile = as.list(quantile(vec, na.rm = T))
  names(l.percentile) = c('min','percentile.25','median','percentile.75','max')
  return(l.percentile)
}
```

**Testcases**:
```{r}
# vanilla case
l.1 = percentile.list(c(1,2,3,4,5))
stopifnot(l.1$min == 1)
stopifnot(l.1$percentile.25 == 2)
stopifnot(l.1$median == 3)
stopifnot(l.1$percentile.75 == 4)
stopifnot(l.1$max == 5)

l.2 = percentile.list(c(4,6,10,20,43,6,8,90,22,1,0,-6,-7,-10))
stopifnot(l.2 == c(-10, 0.25, 6, 17.5, 90))

# NA handling
l.3 = percentile.list(c(1,2,3,NA,4,5,NA))
stopifnot(l.3 == c(1,2,3,4,5))

# empty and all-NA situation
l.4 = percentile.list(c(NA, NA))
l.5 = percentile.list(c())
stopifnot(is.na(l.4) == c(T,T,T,T,T))
stopifnot(is.na(l.5) == c(T,T,T,T,T))
```

3. Write an R function that, when given a vector of strings,
returns the position(s) of the longest strings in the vector.


**Solution**:
```{r}
which.longest = function(str.vec) {
  #' Find the position(s) of longest strings in a vector.
  #' @param str.vec: a vector of strings.
  #' @return the positions.
  return(which(nchar(str.vec)==max(nchar(str.vec), na.rm=T)))
}
```

**Testcases**:
```{r}
# Test cases
# Vanilla case, 3 expected
stopifnot(which.longest(c("a", "ab", "abc", "a")) == c(3))


# Multiple longest strings
# we should find all positions, 2,3,6 expected
stopifnot(which.longest(c("a", "abc", "def", "a", NA, 'ghl')) == 
            c(2, 3, 6))

# NA handling
# 4 expected
stopifnot(which.longest(c(NA, "a", "ab", "abc", NA)) == c(4))

# All-NA vector
# it should not get anything. Actually it gives a warning
# because it's trying to take maximum on all-NA vector. 
# we can live with this behavior, some kind of warning is "expected".
stopifnot(which.longest(c(NA, NA)) == c())

# Empty vector
# it gives a warning, as expected.
stopifnot(which.longest(c()) == c())

# Spaces should be treated like chars
# 2,4 expected
stopifnot(which.longest(c("  ", "     ", "abcd", "abcde")) == c(2,4))
```

4. Create an __infix__ operator that returns whether or not a
number is a multiple of another. Call it %m%. In other words,
`10 %m% 5` should be `TRUE` while `10 %m% 3` should be `FALSE`.

**Solution**:
```{r}
'%m%' = function(vec, divisor) {
  #' calculate whether entries in \code{vec} are multiples 
  #' of \code{divisor}. If \code{vec} has same dimension as
  #' \code{divisor}, the calculation is performed element-wise.
  #' Otherwise, the one with less entries will be duplicated to
  #' match the other's dimension, then perform element-wise calculation.
  #' @param vec: a vector of strings.
  #' @param divisor: a vector of strings.
  #' @return logical vector.
  return(vec %% divisor == 0)
}
```

**Testcases**:
```{r}
# scalar cases
stopifnot(8 %m% 2 == T)   # 8 is multiple of 2
stopifnot(8 %m% 3 == F)   # 8 is not multiple of 3
stopifnot(is.na(8 %m% 0)) # divide 0 is not valid, return NA
stopifnot(is.na(8 %m% NA)) # NA handling
stopifnot(is.na(NA %m% 2)) # NA handling
stopifnot(is.na(NA %m% NA)) # NA handling

# vector cases
# if If vec has same dimension as divisor, the calculation is performed 
# element-wise. Otherwise, the one with less entries will be duplicated to
# match the other's dimension, then perform element-wise calculation.

# case.1, same dimensions
stopifnot(c(2,4,8,16) %m% c(2,2,2,2) == c(T,T,T,T))
stopifnot(c(3,6,9,12) %m% c(2,2,2,2) == c(F,T,F,T))
stopifnot(c(3,6,9,12) %m% c(1,2,3,6) == c(T,T,T,T))
stopifnot((is.na(c(3,6,9,NA) %m% c(NA,2,0,6)) == c(T,F,T,T)))

# case.2, vec has lower dimension
stopifnot(256 %m% c(3,16,77,128) == c(F,T,F,T))
# Right broadcasting, equivalent to c(256, 256, 256, 256) %m% c(...)
stopifnot(c(77,256) %m% c(3,16,77,128) == c(F,T,T,T))
# Right broadcasting, equivalent to c(77, 256, 77, 256) %m% c(...)
stopifnot(c(256,16,77) %m% c(3,16,77,128) == c(F,T,T,T))
# Right broadcasting, equivalent to c(256, 16, 77, 256) %m% c(...)
# a warning is generated saying longer length is not multiple of the shorter length
# so broadcasting is partially done.

# case.3, divisor has lower dimension
stopifnot(c(2,4,8,16) %m% 8 == c(F,F,T,T))
# Left broadcasting, equivalent to c(...) %m% c(8, 8, 8, 8)
stopifnot(c(2,4,8,16) %m% c(2,8) == c(T,F,T,T))
# Left broadcasting, equivalent to c(...) %m% c(2, 8, 2, 8)
stopifnot(c(2,4,8,16) %m% c(16,8,4) == c(F,F,T,T))
# Left broadcasting, equivalent to c(...) %m% c(16, 8, 4, 16)
# a warning is generated saying longer length is not multiple of the shorter length
# so broadcasting is partially done.
```


5. Go to the following website:
[https://www.sec.gov/data/foiadocsfailsdatahtm](https://www.sec.gov/data/foiadocsfailsdatahtm)
and download the __July 2017, Second Half__ data set.
Read these data into R and create an appropriate data frame.
Be sure that each column is in an appropriate form (date, factor,
character, etc.)
When reading in the file, __do not make any changes to the data
file itself__. Resolve any issues using appropriate R commands.


**Issues**:



- The data is parsed with delimiter `|`.
- The last two rows are not part of data, we should skip them.
- Several columns do not have correct type. In particular:
    - `SETTLEMENT.DATE` should be date field.
    - `DESCRIPTION` had better be a char field.
    - `SYMBOL` and `CUSIP` are tickers symbols and CUSIP numbers. They may serve as keys to query the data, so we leave them as factor fields.
    - `PRICE` should be a numeric field.
    
    
**Solution**:
```{r}
df = read.delim('cnsfails201707b.txt',sep='|',
                nrow=length(readLines("cnsfails201707b.txt")) - 3)
str(df)
```

```{r}
df$SETTLEMENT.DATE = as.Date.character(df$SETTLEMENT.DATE, format='%Y%m%d')
df$DESCRIPTION = as.character(df$DESCRIPTION)
df$PRICE = as.character(df$PRICE)
df$PRICE = replace(as.character(df$PRICE), df$PRICE==".","0.00")
df$PRICE = as.numeric(df$PRICE)
str(df)
```

