---
title: "Homework 6"
author: "Ze Yang (zey@andrew.cmu.edu)"
date: "Due Thursday, October 12 at 3:00 PM"
output: pdf_document
---

\large

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(quantmod)
```

You should submit the Rmd file for your analysis.
Name the file as `YOURANDREWID_HW6.Rmd`
and submit it via Canvas. Also submit the `.pdf` file that
is produced.

\vspace{.2in}

__Part 1:__

Suppose we are interested in building a training set of features that
we hope could be useful in a prediction problem. In particular, the
idea we want to explore is the use of the percentiles of the daily closing
stock price over the past $n$ weeks.

Write R code to perform this task.

Use smoothing to estimate the percentiles. You can use the code that was
provided and described in lecture.

The sample should be constructed by randomly sampling NYSE ticker symbols.
An R package/function for obtaining the list of all NYSE ticker symbols
was described in lecture.

Market data can be read in using the `quantmod` package. Again, this was
presented in lecture.

Your code should assume that $n$ is a positive integer, and $n \geq 2$. The following example illustrates how easy `quantmod` makes it to
get the $n$ most recent weeks of data:

```{r, warning=FALSE, message=FALSE, eval=FALSE}
foo = getSymbols("AAPL", from=Sys.Date()-n*7, to=Sys.Date(), 
                 auto.assign=FALSE)
```

Choose the bandwidth for the kernel density estimators using the Sheather-Jones
approach.

Use the "Adjusted" price to calculate the quantiles.

Sample 100 NYSE ticker symbols at random for this purpose.

The final result should be a data frame. Each row should correspond to a different
ticker symbol. The first column should be filled with the ticker symbol. The following columns
should be the percentiles. For this exercise, estimate
the $10^{th}$, $20^{th}$, $\ldots$, $90^{th}$ percentiles. Hence, the data
frame should have 100 rows and ten columns.


\vspace{.2in}

**Solution:**


```{r}
stock.sample = function(tickers, n_weeks, 
                        n_samples=100, seed=42, ...) {
  #' Download adjusted close prices for sample tickers.
  #' @param tickers: vector, the whole tickers set.
  #' @param n_weeks: int, number of weeks to look at.
  #' @param n_samples: int, number of tickers to draw.
  #' @param seed: int, the random seed.
  #' @return data.frame of prices, where each row is a date 
  #' and each column is a ticker.
  require(quantmod)
  set.seed(seed)
  tickers.shuffle = sample(tickers)
  # i: current ticker, i.got: number of good tickers.
  # (good: not 404 error, more than 1 rows)
  i = 1; i.got = 1
  price.list = list()
  downloaded.tickers = c()
  while (i.got <= n_samples && i <= length(tickers)) {
    out = tryCatch({
      # fetch stock data.
      df = getSymbols(
        tickers.shuffle[i], 
        from=Sys.Date()-n_weeks*7, to=Sys.Date(), 
        auto.assign=FALSE)
      # raise error(continue) 
      # if the downloaded column has only 1 row.
      if (length(df[,1]) == 1) { 
        stop("Abnormal data: Only one row.\n") 
      }
      # increment i.got only if getting a "good" ticker.
      i.got = i.got + 1
      downloaded.tickers = c(
        downloaded.tickers, tickers.shuffle[i])
      Ad(df)
    },
    # error handling: continue
    error = function(e){message(e)},
    # warning handling: continue
    warning = function(e){message(e)})
    price.list[[tickers.shuffle[i]]] = out
    # increment i
    i = i+1
    # print progress
    if (i.got %% 10 == 0) {
      print(sprintf(
        "[quantmod] %d tickers downloaded.", i.got))
    }
  }
  return(price.list)
}

kde.cdf = function(x, bw='SJ', res=100, ...) {
  #'Get the cdf and inverse cdf of a kde distribution.
  #'@param x: vector, the data set over which the kde is computed.
  #'@param bw: string, the method used to compute kde binwidth.
  #'@return List of two functions, the linear interpolated
  #'cdf and inverse cdf of the kernel density estimation.
  kde = density(x, bw=bw, ...)
  interp.kde = approxfun(
    kde$x, kde$y, yleft=0, yright=0)
  lb = min(kde$x)
  ub = max(kde$x)
  x.grid = seq(lb, ub, length=res)
  cdf.grid = numeric(res)
  for (i in 1:res) {
    cdf.grid[i] = integrate(
      interp.kde, lower=lb, upper=x.grid[i], 
      stop.on.error=F)$value
  }
  cdf.fun = approxfun(
    x.grid, cdf.grid, yleft=0, yright=1)
  invcdf.fun = approxfun(
    cdf.grid, x.grid, yleft=NA, yright=NA)
  return(list(cdf=cdf.fun, inv.cdf=invcdf.fun))
}

stock.kde.quantiles = function(
  tickers, probs=c(0, 0.25, 0.5, 0.75, 1), n_weeks=2, 
  n_samples=100, bw='SJ', res=100, seed=42) {
  #' Wrapper function of stock.sample and kde.cdf.
  #' @param tickers: vector, the whole tickers set.
  #' @param probs: vector, the prob level of quantiles to calculate.
  #' @param n_weeks: int, number of weeks to look at.
  #' @param n_samples: int, number of tickers to draw.
  #' @return data.frame, where each row is a ticker, 
  #' and each column is a probability level.
  for (p in probs) {
    if (p > 1 || p < 0) {stop("Invalid probability.")}
  }
  quantiles.list = list()
  prices = stock.sample(tickers, n_weeks, n_samples, seed)
  ticker.sample = names(prices)
  for (t in ticker.sample) {
    cdf.holdout = kde.cdf(prices[[t]], bw, res)
    inv.cdf = cdf.holdout$inv.cdf
    quantiles.list[[t]] = inv.cdf(probs)
  }
  df = t(data.frame(quantiles.list))
  colnames(df) = paste(probs, '.quantile', sep='')
  return(df)
}
```


```{r, warning=FALSE} 
# Get all NYSE symbols
tickers = stockSymbols()
NYSE = tickers[tickers$Exchange=='NYSE','Symbol']
probs = seq(0, 0.9, len=10)

qtl.df = stock.kde.quantiles(NYSE, probs, n_weeks=5)
str(qtl.df)
head(qtl.df)
```


---


__Part 2:__

Read in the options sample that I presented in lecture. __(Read in the sample I created,
do not generate your own.)__

Create a plot that compares (1) the difference between the Black-Scholes price and the
ask price with (2) the time to expiration. Smooth the relationship using local linear
regression, and show this on the plot. Is there evidence of a relationship between
these two quantities? Can you guess as to why you are seeing this?

**Solution:**

```{r}
# read option sample
options.data = read.table("optionssample09302017.txt", sep=',', header=T)
# difference between Black-Scholes price and market price.
options.data$price.diff = options.data$bsval  - options.data$ask

ggplot(options.data, aes(x=timetoexpiry, y=price.diff)) +
  geom_point(size=0.5) + geom_smooth(
    method = "loess", method.args=list(degree=1)) +
  labs(x='time to maturity', y='bs price - ask')
```


**Comments:**

- There is no significant pattern between (1) the difference between Black-Scholes price and the market ask price and (2) the time to maturity of the option. The slope of tge fitted local linear regression line is almost zero, and the price differences are just scattered around the line very randomly.

- As a result, we may conclude that the Black-Scholes price matches with the real market price very well on the $T$ (time to maturity) dimension.

- $T$ is an input to Black-Scholes model. It is likely that the variations in option prices that attributes to $T$ is well captured by the Black-Scholes model. The remaining errors are caused by other factors that are not included in BS model. So the errors have little correlation with $T$, and hence are identically distributed for different $T$'s.
